---
title: "Case Study 1 Initial"
author: "Jim Holdnack"
date: "February 23, 2019"
output:
  html_document:
    keep_md: yes
  keep_md: default
---


### XYZ Brewery Porposal
According to Fortune.com1: US Craft Brew industry is worth $26 billion dollars in 2017
This represents increases of 6.2 billion since 2015
The pace of growth is slowing but the opportunity is large
Craft beer drinkers tend to support local independent breweries
Consumers are becoming more selective
Rising demand for low alcohol by volume (ABV) and flavored beer
Each market share of .5% = $130 million in revenue

STEP 1: "Competitive Landscape" Evaluate current production of craft brews regionally
How many craft breweries are currently producing in the US?



```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
## obtain data needed to evaluate competitive landscape and load into R
getwd()
setwd("c:/Users/jhold/Desktop/case study 1")
library(tidyverse)
beers <- read_csv("c:/Users/jhold/Desktop/case study 1/Beers.csv")
brewers <- read_csv("c:/Users/jhold/Desktop/case study 1/Breweries.csv")
#check content of files before analysis
str(beers)
str(brewers)
# How many breweries are in the US
length(brewers$State)



```
### Combine multiple data sources to provide richer understanding of craft brewering in US
Check integrity of data files
evaluate missing data

```{r, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
# fully analyze data by combining different data sets
Brewtot <- merge(beers, brewers, by.x = "Brewery_id", by.y = "Brew_ID")
# check integrity of combined files and evaluate variable features
head(Brewtot,6)
tail(Brewtot,6)
str(Brewtot)
### Determine market saturation by state for breweries and beers brewed
library(plyr)
BrewersbyState <- brewers$State
CntState <- count(brewers$State)
cntbeers <- count(Brewtot$State)
df <- merge(CntState,cntbeers, by.x ='x', by.y='x')
names(df) <- c("State", "Breweries", "Unique Brews")
library (openintro)
library(gridExtra)
library(grid)
df$StateName <- abbr2state(df$State) 
df <- df[order(-df$Breweries),] 
dft <- df[,c(4,2,3)]
grid.table(dft,rows = NULL)
```
### Tables must be separately generated
```{r, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

dfh <-df[1:10,c(4,2,3)]
grid.table(dfh,rows = NULL)
```

```{r, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
dfl <-df[42:51,c(4,2,3)]  
grid.table(dfl,rows=NULL)
```

# Evaluate for missing data points


```{r, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

MissTot <- sapply(Brewtot, function(x) sum(is.na(x)))
MissTot
# IBU's have a high rate of missing data which may effect analyses of these data points

```

## Summary of Brewery data
The Top 5 States for number of breweries are:
Colorado = 42,  California=  39, Michigan= 32, Oregon= 29, Texas= 28
States with only 1 craft brewery are:
District of Columbia, North Dakota, South Dakota, and West Virginia
There are 2410 unique craft beers brewed across the US
The Top 5 States for number of unique beers brewed are:
Colorado= 265, California= 183, Michigan= 162, Indiana=139, Texas= 130
The lowest 5 States for number of unique beers brewed are:
Tennessee= 6, Arizona=5, North Dakota=3, Delaware=2, West Virginia=2



## Are there regional differences in beer characteristics
Median Alcohol by Volume (ABV) and median International Bitterness Unit (IBU) inform about the average characteristics of craft beers
What is the distribution of ABV and IBU and what does that tell us about beer brewing in the US market
What is the median ABV for each state?
What is the median IBU for each state?


```{r, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

MedA<- aggregate (Brewtot$ABV, list(Brewtot$State), median, na.rm=TRUE)
MedI <- aggregate (Brewtot$IBU, list(Brewtot$State), median, na.rm=TRUE)

MedA
MedI
#### South Dakota is missing IBU data and will be dropped as a data point
MedI <- MedI[-c(42),]
## evaluate the frequency distributions of ABV and IBU 
hist(Brewtot$ABV, main="Distribution of ABV", breaks=20, xlab="Alcohol by Volume", border="black", col="blue", xlim=c(.02,.13)) 
hist(Brewtot$IBU, main="Distribution of IBU", breaks=20, xlab="International Bitterness Units", border="black", col="red", xlim=c(0,100))
## Both ABV and IBU's are positively skewed. Most brews have a low to moderate IBU and ABV while a subset of Beers are high in ABV and/or IBU
## Evaluate beer characteristics by state to determine if thwere any regional effects to consider
ggplot(data=MedA, aes(x=reorder(Group.1, x), y=x, fill=Group.1)) +geom_bar(stat="identity") +coord_flip() +ylab("Median ABV") + xlab("State Name") +ggtitle("Median ABV by State")+ theme(plot.title = element_text(hjust = 0.5)) + theme(legend.position="none")
ggplot(data=MedI, aes(x=reorder(Group.1, x), y=x, fill=Group.1)) +geom_bar(stat="identity") +coord_flip() +ylab("Median IBU") + xlab("State Name") +ggtitle("Median IBU by State")+ theme(plot.title = element_text(hjust = 0.5)) + theme(legend.position="none")
## Evaluate alcohol level
Brewtot <- Brewtot[order(-Brewtot$ABV),] 
##Most alcohol by volume beer is Colorado
head(Brewtot,1)
## highest 5 and lowest 5 state median abv's
MedA <- MedA[order(-MedA$x),] 
names(MedA) <- c("State", "ABV")
head(MedA,5)
tail(MedA,5)
## basic statistics for ABV use pastecs for standard deviation and to check standard error of the mean
library(pastecs)
options(scipen=100)
stat.desc(Brewtot$ABV)
### Evaluate IBU level
Brewtot <- Brewtot[order(-Brewtot$IBU),] 
## Most bitter beer by IBU is brewed in 
head(Brewtot,1)
## highest 5 and lowest 5 state median IBU's
MedI <- MedI[order(-MedI$x),] 
names(MedI) <- c("State", "IBU")
head(MedI,5)
tail(MedI,5)
stat.desc(Brewtot$IBU)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


#### Beer Characteristcs overall and by State
### ABV Data
Highest ABV reported = .128 
Lee Hill Series Vol. 5 - Belgian Style Quadrupel, Upslope Brewing Company      Boulder, CO
Overall ABV Characteristics
Mean= 0.0598, Median=.056, Standard Deviation=0135
States with Highest Average (median) ABV:
District of Columbia=.0625, Kentucky=.0625, Michigan=.062,
   New Mexico=.62, and West Virginia=.62
States with Lowest Average (median) ABV:
Kansas=.05, North Dakota=.05, Wyoming=.05, New Jersey=.046, and Utah=.04

### IBU Data
Highest IBU reported = 138 
Bitter Bitch Imperial IPA, American Double / Imperial IPA, Astoria Brewing Company Astoria, OR
Mean=42.7, Median=35.0, Standard Deviation=25.9
States with Highest Average (median) IBU:
Maine=61.0, West Virginia=57.5, Florida=55.0, Georgia=55.0, Delaware=52.0
States with Lowest Average (median) IBU:
Hawaii=22.5, Wyoming=21.0, Arizona=20.5, Kansas=20.0, and Wisconsin=19.0


#### what is the relationship between ABV and IBU

```{r, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

### plot of ABV by IBU
ggplot(data=Brewtot, aes(x=IBU, y=ABV)) +geom_point(shape = 16, size = 2, color="blue") + stat_smooth(method = 'lm', color='red') + labs(title = 'Craft Beer IBU by ABV') + theme(plot.title = element_text(hjust = 0.5))
### correlation test of limear relationship
cor.test( ~ ABV + IBU,data=Brewtot,method = "pearson")


```


#### Plot of IBU versus ABV
Shows a general positive trend with higher Abv associated with higher levels of IBU
The regression line illustrates that the variability may be changing along the prediction line (heteroscedasticity)
    This is exemplified by the fact that beers of 5% ABV or lower tend to have low to moderate IBU and not high levels of IBU
    However, at ABV of 7.5 and higher, you get a full range of IBU's meaning that it is possible to have high ABV low IBU beers
    Note: this coreelation does not show causation as the relationship may be affect by third variables such as ingredients used, 
     fermentaiton process, and marketing needs
     
### Data regarding consumption is helpful to make business plans
additional data was scraped from URL=http://scottjanish.com/map-per-capita-gallons-beer-consumed-per-adult-state


```{r, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
### scraping web data on consumption
library(XML)
library(rvest)
dr1=html('http://scottjanish.com/map-per-capita-gallons-beer-consumed-per-adult-state')
drinks  <- dr1 %>%  html_nodes("td") %>% html_text()
str(drinks)
x <- seq(1,101,2)
States <- drinks[x]
y <- seq(2,102,2)
Consumption <- as.numeric(drinks[y])
dr2<-dplyr::data_frame(States,Consumption)
str(dr2)
dr2$State <- state2abbr(dr2$States) 
dr2 <- dr2[order(-dr2$Consumption),] 
head(dr2)
tail(dr2)
drh <- dr2[1:10,c(1,2)]
drl <- dr2[42:51,c(1,2)]

```

#### Regions with higher consumption rates fall in the least saturated category (ND,SD)
Montana is regionally close. 
Competition from Colorado & Wisconsin

```{r, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
grid.table(drh,rows = NULL)
```
#### Consumption in gallons per adult
```{r, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
grid.table(drl,rows=NULL)
```
#### Consumption in gallons per adult
